---
title: "dplyr"
author: "Sarah Lee"
date: "1/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Resources

* Jenny Bryan's STAT545: [Introduction to dplyr](https://stat545.com/dplyr-intro.html#think-before-you-create-excerpts-of-your-data)
* Hadley Wickham & Garrett Grolemond's R for Data Science: [Data Transformation](https://r4ds.had.co.nz/transform.html#introduction-2)
* Ocean Health Index: [Data Wrangling](http://ohi-science.org/data-science-training/dplyr.html)


## Getting Started

Load the `tidyverse` package, which includes `dplyr`, and the `gapminder` package with data

```{r load tidyverse}
library(tidyverse)  ## install.packages("tidyverse")
```

```{r load gapminder}
library(gapminder)  ## install.packages("gapminder")
```

## The gapminder data

```{r}
gapminder
```

## Comparison operators in logical expressions

*True or False*: The object y less than or equal to 15.

```{r}
y <- 9  ## assign the value of 9 to the object y

y <= 15 ## create a logical expression that will return either TRUE or FALSE
```

'filter' takes logical expressions and returns the rows for which all are `TRUE`

```{r}
filter(gapminder, lifeExp < 29)
```

## Mind your equals signs

Common error: using = instead of ==

```{r}
# filter(gapminder, year = 1980)
#>Error: `year` (`year = 1980`) must not be named, do you need `==`?

filter(gapminder, year ==1967)
```

## Logical operators

Used to combine multiple arguments.

 logical expressions and returns the rows for which all are `TRUE`

You will also need to load the package '

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
